---
title: "Data Wrangling 1 _ Data Manipulation"
author: "Jiayi Shen"
date: "9/20/2018"
output: html_document
---

```{r, include = FALSE}
#knitr::opts_chunk$set
library(tidyverse)
```

```{r}
options(tibble.print_min = 3)

litters_data = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv("./data/FAS_pups.csv",
  col_types = "ciiiii")
pups_data = janitor::clean_names(pups_data)

```


## Select variables

```{r}
select(litters_data, group, litter_number, gd0_weight, pups_born_alive)
```


... by specifying a range:
```{r}
select(litters_data, group:gd_of_birth)
select(litters_data, group, gd_of_birth:pups_survive)
```


... by saing what you want to remove:
```{r}
select(litters_data, -litter_number)
```

select and rename in the same process:
```{r}
select(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number)
rename(litters_data, LiTtEr_NuMbEr = litter_number)
```

... use select helpers:
```{r}
select(litters_data, starts_with("gd"))
select(litters_data, litter_number, starts_with("pup"))

# moving one coloum to the left and everything else follows.
select(litters_data, litter_number, everything())
```

in-class assessment 1:
```{r}
select(pups_data, litter_number, sex, pd_ears)
```



## Filter

```{r}
filter(litters_data, gd0_weight < 25)
filter(litters_data, gd0_weight >= 25)
filter(litters_data, pups_born_alive == 8)

filter(litters_data, is.na(gd0_weight))
filter(litters_data, !is.na(gd0_weight))

# using the "|"(or) operator
filter(litters_data, group == "Low8" | group == "Low7")


filter(litters_data, pups_born_alive %in% 4:6)
filter(litters_data, pups_born_alive %in% 4:6, !is.na(gd0_weight))
filter(litters_data, pups_born_alive %in% 4:6 & !is.na(gd0_weight))
```

in-class assessment 2:
```{r}
filter(pups_data, sex == 1)
filter(pups_data, pd_walk < 11, sex == 2)
```

## Mutate
create varialbes 

```{r}
mutate(litters_data,
       wt_gain = gd18_weight - gd0_weight,#create new variable based on old varialbes
       group = tolower(group) #tolower makes everything lower case. 
       )

mutate(litters_data,
       wt_gain = gd18_weight - gd0_weight,
       wt_gain_squared = wt_gain^2
      )
```

## Arrange
```{r}
arrange(litters_data, gd0_weight)

#organizing in the descending order
arrange(litters_data, pups_born_alive, desc(gd0_weight))
```


## Piping

look at intermediate object approach:
```{r}
litters_data_raw = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data_clean_names = janitor::clean_names(litters_data_raw)
litters_data_selected_cols = select(litters_data_clean_names, -pups_survive)
litters_data_with_vars = mutate(litters_data_selected_cols, 
  wt_gain = gd18_weight - gd0_weight,
  group = tolower(group))
litters_data_with_vars
```
intermediate object approach takes up unnecessary working space.\
\
\
now use nested function calls:
```{r}
litters_data_clean = 
  mutate(
    select(
      janitor::clean_names(
        read_csv("./data/FAS_litters.csv", col_types = "ccddiiii")
        ), 
    -pups_survive
    ),
  wt_gain = gd18_weight - gd0_weight,
  group = tolower(group)
  )
litters_data_clean
```
Same data processing as the previous approach, but neither of these are easy to read.


Use Piping!!
hot key for "%>%" : shift+command+M
```{r}
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names(dat = .) %>%
  select(.data = ., -pups_survive) %>%
  mutate(.data = .,
    wt_gain = gd18_weight - gd0_weight,
    group = tolower(group))
```

placeholder . 
lm : linear holder function
```{r}
processed_litters_data =   
read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names(dat = .) %>%
  select(.data = ., -pups_survive) %>%
  mutate(.data = .,
    wt_gain = gd18_weight - gd0_weight,
    group = tolower(group)) %>% 
  lm( gd0_weight ~ gd18_weight, data = .)
```

